---
title: "  **Chi-Square and ANOVA Tests**"
output: pdf_document
header-includes:
  - \usepackage{fancyhdr}
  - \usepackage{setspace}\spacing{1.5}
  - \usepackage{titling}
  - \pretitle{\begin{center}
    \includegraphics[width=6in,height=11in]{image.png}\LARGE\\}
  - \posttitle{\end{center}}
  - \usepackage{sectsty} \sectionfont{\emph}
---
\begin{centering}
\vspace{1cm}
{\bf Submitted By : David Joseph Johnson } \\
\vspace{.75cm}
{\bf 01/23/2022 } \\
\end{centering}
\newpage
\tableofcontents
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Hypothesis Testing is a method used to test an assumption about the entire population by picking a random sample.Two likely conflicting hypotheses are stated (Null Hypothesis H0 and Alternative Hypothesis H1), and based on the inference drawn from the sample, one of the hypothesis is accepted.

**Chi-square Test** is a hypothesis testing method.The most common two Chi-Square tests are:

* Chi-Square Goodness of Fit Test

* Chi-Square Test of Independence

The goodness of fit is used if a single measurement variable is being analyzed to determine likely distribution.The test of independence is used to validate if the two measurement variables are related or not.

**ANOVA (Analysis of Variance) Test** are used to find the significance of the survey or result of an experiment.ANOVA tests helps overcome the issue of only being able analyze two variables in the T-test by considering more than two variables.In other words, multiple groups can be considered and the difference between them can be tested.ANOVA is a special regression where the predictors are all categorical.

* One-Way ANOVA has one independent variable

* Two-Way ANOVA has two independent variable

*In this assignment Chi-square and ANOVA tests are used to solve tasks using R.*

\newpage

# Analysis

```{r , echo=FALSE ,results='hide', warning=FALSE, message=FALSE}
### Invoke Packages
library("ggplot2")
library("corrplot")
library("dplyr")
library("janitor")
library("tidyr")
library("RColorBrewer")
library("car")
library("psych")
library("imputeTS")
library("corrplot")
library("MASS")
library("leaps")
library("knitr")
library("ggplot2")
library("tibble")
```

## Task 1.(Section 11-1 6. Blood Types)  

A medical researcher wishes to see if hospital patients in a large hospital have the same blood type distribution as those in the general population. The distribution for the general population is as follows: type A, 20\%; type B, 28\%; type O, 36\%; and type AB = 16\%. He selects a random sample of 50 patients and finds the following: 12 have type A blood, 8 have type B, 24 have type O, and 6 have type AB blood.
At alpha = 0.10, can it be concluded that the distribution is the same as that of the general population?

```{r }
# Method:  Chi-Square Goodness of Fit Test
# 1.a (Hypotheses)
# H0: A = 20%, B = 28%, O = 36%, AB = 16%
# H1: Distribution is not as stated in null hypothesis

# Set significance level
alpha <- 0.10
# Vector of probabilities of general population
prob <- c(0.20, 0.28, 0.36, 0.16)
# Vector of observed values of sample
observed <- c(12, 8, 24, 6)

result <- chisq.test(x= observed, p= prob)
# Test Statistic
result
# 1.b (Critical Value)
qchisq(p=alpha, df=result$parameter, lower.tail=FALSE)
# 1.c (Test Value)
result$statistic 
# 1.d
# Compare p-value and alpha to make decision
ifelse(result$p.value > alpha,"Failed to reject Null Hypothesis","Reject Null Hypothesis")
# 1.e
# Sample Distribution is same as population
```


## Task 2.(Section 11-1. 8. On-Time Performance by Airlines) 

According to the Bureau of Transportation Statistics, on-time performance by the airlines is described as follows: 
```{r , echo=FALSE , warning=FALSE, message=FALSE}
df <- data.frame(Action = c("On time","National Aviation System delay","Aircraft arriving late","Other (because of weather and other conditions)"),
                Percentage_of_Time  = c(70.8,8.2,9.0,12.0))

print(kable(df))
```

Records of 200 randomly selected flights for a major airline company showed that 125 planes 
were on time; 40 were delayed because of weather, 10 because of a National Aviation System 
delay, and the rest because of arriving late. At alpha = 0.05, do these results differ from the 
government’s statistics? 

```{r }
# Method:  Chi-Square Goodness of Fit Test
# 2.a (Hypotheses)
# H0: On time = 70.8%, National Aviation System delay = 8.2%, Aircraft arriving late = 9%,
#     Other (because of weather and other conditions) = 12%
# H1: Distribution is not as stated in null hypothesis

# Set significance level
alpha <- 0.05
# Vector of probabilities of general population
prob <- c(0.708, 0.082, 0.09, 0.12)
# Vector of observed values of sample
observed <- c(125, 10, 25, 40)

result <- chisq.test(x= observed, p= prob)
# Test Statistic
result
# 2.b (Critical Value)
qchisq(p=alpha, df=result$parameter, lower.tail=FALSE)
# 2.c (Test Value)
result$statistic 
# 2.d
# Compare p-value and alpha to make decision
ifelse(result$p.value > alpha,"Failed to reject Null Hypothesis","Reject Null Hypothesis")
# 2.e
# Sample Distribution is not same as population
```


## Task 3.(Section 11-2.8. Ethnicity and Movie Admissions )  

Are movie admissions related to ethnicity? A 2014 study indicated the following numbers of admissions (in thousands) for two different years. At the 0.05 level of significance, can it be concluded that movie attendance by year was dependent upon ethnicity?  

```{r , echo=FALSE , warning=FALSE, message=FALSE}
df <- data.frame(
              Year  = c(2013,2014),
                Caucasian  = c(724,370),
                Hispanic  = c(335,292),
                African_American  = c(174,152),
                Other  = c(107,140))

print(kable(df))
```

```{r }
# Method:  Chi-Square of Independence
# 3.a (Hypotheses)
# H0: Movie admissions are independent of ethnicity
# H1: Movie admissions are dependent of ethnicity

# Set significance level
alpha <- 0.05
# Creating Vector for each row
r1 <- c(724,335,174,107)
r2 <- c(370,292,152,140)
# Vector of observed values of sample
data_matrix = matrix(c(r1,r2),nrow = 2, byrow = TRUE)
# Name rows and columns of matrix
rownames(data_matrix) = c("2013","2014")
colnames(data_matrix) = c("Caucasian","Hispanic","African_American","Other")
# Resultant Matrix
data_matrix

# Chi-Sq Test
result <- chisq.test(data_matrix)
# Test Statistic
result
# 3.b (Critical Value)
qchisq(p=alpha, df=result$parameter, lower.tail=FALSE)
# 3.c (Test Value)
result$statistic 
# 3.d
# Compare p-value and alpha to make decision
ifelse(result$p.value > alpha,"Failed to reject Null Hypothesis","Reject Null Hypothesis")
# 3.e
# Movie admissions are dependent of ethnicity
```

\newpage
## Task 4.(Section 11-2.10 Women in the Military)  

This table lists the numbers of officers and enlisted personnel for women in the military. At alpha = 0.05, is there sufficient evidence to conclude that a relationship exists between rank and branch of the Armed Forces?  
```{r , echo=FALSE , warning=FALSE, message=FALSE}
df <- data.frame(Department  = c("Army","Navy","Marine Corps","Air Force"),
                Officers  = c(10791,7816,932,11819),
                Enlisted  = c(62491,42750,9525,54344))

print(kable(df))
```

```{r }
# Method:  Chi-Square of Independence
# 4.a (Hypotheses)
# H0: Ranks of women in military are independent of branch
# H1: Ranks of women in military are dependent of branch

# Set significance level
alpha <- 0.05
# Creating Vector for each row
r1 <- c(10791,62491)
r2 <- c(7816,42750)
r3 <- c(932,9525)
r4 <- c(11819,54344)
# Vector of observed values of sample
data_matrix = matrix(c(r1,r2,r3,r4),nrow = 4, byrow = TRUE)
# Name rows and columns of matrix
rownames(data_matrix) = c("Army","Navy","Marine Corps","Air Force")
colnames(data_matrix) = c("Officers","Enlisted")
# Resultant Matrix
data_matrix

# Chi-Sq Test
result <- chisq.test(data_matrix)
# Test Statistic
result
# 4.b (Critical Value)
qchisq(p=alpha, df=result$parameter, lower.tail=FALSE)
# 4.c (Test Value)
result$statistic 
# 4.d
# Compare p-value and alpha to make decision
ifelse(result$p.value > alpha,"Failed to reject Null Hypothesis","Reject Null Hypothesis")
# 4.e
# Ranks of women in military are dependent of branch
```
\newpage
## Task 5.(Section 12-1.8 Sodium Contents of Foods)   

The amount of sodium (in milligrams) in one serving for a random sample of three different kinds of foods is listed. At the 0.05 level of significance, is there sufficient evidence to conclude that a difference in mean sodium amounts exists among condiments, cereals, and desserts?   
```{r , echo=FALSE , warning=FALSE, message=FALSE}
df <- data.frame(Condiments  = c(270,130,230,180,80,70,200,0),
                Cereals  = c(260,220,290,290,200,320,140,0),
                Desserts  = c(100,180,250,250,300,360,300,160))
print(kable(df))
```
 *Note: Zero values are not considered as individual data value*

```{r }
# Method:  One-way ANOVA Test
# 5.a (Hypotheses)
# H0: mean(Condiments) = mean(Cereals) = mean(Desserts)
# H1: Atleast one mean is different from others

# Set significance level
alpha <- 0.05
# Dataframe for Condiments
condiments <- data.frame('sodium' = c(270,130,230,180,80,70,200),
                         'food' = rep('condiments',7), stringsAsFactors = FALSE)
# Dataframe for Cereals
cereals <- data.frame('sodium' = c(260,220,290,290,200,320,140),
                      'food' = rep('cereals',7), stringsAsFactors = FALSE)
# Dataframe for Desserts
desserts <- data.frame('sodium' = c(100,180,250,250,300,360,300,160),
                       'food' = rep('desserts',8), stringsAsFactors = FALSE)
# Combine the Dataframe
sodium <- rbind(condiments,cereals,desserts)
sodium$food <- as.factor(sodium$food)

# ANOVA Test
anova <- aov(sodium ~ food, data = sodium)

# Model Summary
a_summ <- summary(anova)
a_summ
# 5.b (Critical Value)
qf(1-alpha,a_summ[[1]][1,1],a_summ[[1]][2,1])

# 5.c (Test Value)
F.value <- a_summ[[1]][[1,"F value"]]
F.value
# 5.d
# Compare p-value and alpha to make decision
p.value <- a_summ[[1]][[1,"Pr(>F)"]]
p.value
ifelse(result$p.value > alpha,"Failed to reject Null Hypothesis","Reject Null Hypothesis")
# 5.e
# Atleast one mean is different from others
```


## Task 6.(Section 12-2.10 Sales for Leading Companies)    

The sales in millions of dollars for a year of a sample of leading companies are shown. At alpha = 0.01, is there a significant difference in the means?    
```{r , echo=FALSE , warning=FALSE, message=FALSE}
print("Chocolate")
df <- data.frame(Cereals  = c(578,320,264,249,237),
                Candy  = c(311,106,109,125,173),
                Coffee  = c(261,185,302,689,0))

print(kable(df))
```
*Note: Zero values are not considered as individual data value*

```{r }
# Method:  One-way ANOVA Test
# 6.a (Hypotheses)
# H0: mean(Cereals) = mean(Candy) = mean(Coffee)
# H1: Atleast one mean is different from others

# Set significance level
alpha <- 0.01
# Dataframe for Cereals
cereals <- data.frame('sales' = c(578,320,264,249,237),
                      'chocolate' = rep('cereals',5), stringsAsFactors = FALSE)
# Dataframe for Candy
candy <- data.frame('sales' = c(311,106,109,125,173),
                    'chocolate' = rep('candy',5), stringsAsFactors = FALSE)
# Dataframe for Coffee
coffee <- data.frame('sales' = c(261,185,302,689),
                     'chocolate' = rep('coffee',4), stringsAsFactors = FALSE)
# Combine the Dataframe
sales <- rbind(cereals,candy,coffee)
sales$chocolate <- as.factor(sales$chocolate)

# ANOVA Test
anova <- aov(sales ~ chocolate, data = sales)

# Model Summary
a_summ <- summary(anova)
a_summ
# 6.b (Critical Value)
qf(1-alpha,a_summ[[1]][1,1],a_summ[[1]][2,1])
# 6.c (Test Value)
F.value <- a_summ[[1]][[1,"F value"]]
F.value
# 6.d
# Compare p-value and alpha to make decision
p.value <- a_summ[[1]][[1,"Pr(>F)"]]
p.value
ifelse(result$p.value > alpha,"Failed to reject Null Hypothesis","Reject Null Hypothesis")
# 6.e
# Atleast one mean is different from others
```


## Task 7.(Section 12-2.12 Per-Pupil Expenditures)    

The expenditures (in dollars) per pupil for states in three sections of the country are listed. Using alpha = 0.05, can you conclude that there is a difference in means?     
```{r , echo=FALSE , warning=FALSE, message=FALSE}
df <- data.frame(Eastern_third  = c(4946,5953,6202,7243,6113),
                Middle_third  = c(6149,7451,6000,6479,0),
                Western_third  = c(5282,8605,6528,6911,0))

print(kable(df))
```
*Note: Zero values are not considered as individual data value*

```{r }
# Method:  One-way ANOVA Test
# 7.a (Hypotheses)
# H0: mean(Eastern_third) = mean(Middle_third) = mean(Western_third)
# H1: Atleast one mean is different from others

# Set significance level
alpha <- 0.05
# Dataframe for Eastern_third
eastern_third <- data.frame('expenditure' = c(4946,5953,6202,7243,6113),
                      'region' = rep('Eastern_third',5), stringsAsFactors = FALSE)
# Dataframe for Middle_third
middle_third <- data.frame('expenditure' = c(6149,7451,6000,6479),
                    'region' = rep('Middle_third',4), stringsAsFactors = FALSE)
# Dataframe for Western_third
western_third <- data.frame('expenditure' = c(5282,8605,6528,6911),
                     'region' = rep('Western_third',4), stringsAsFactors = FALSE)
# Combine the Dataframe
expenditure <- rbind(eastern_third,middle_third,western_third)
expenditure$region <- as.factor(expenditure$region)

# ANOVA Test
anova <- aov(expenditure ~ region, data = expenditure)

# Model Summary
a_summ <- summary(anova)
a_summ
# 7.b (Critical Value)
qf(1-alpha,a_summ[[1]][1,1],a_summ[[1]][2,1])
# 7.c (Test Value)
F.value <- a_summ[[1]][[1,"F value"]]
F.value
# 7.d
# Compare p-value and alpha to make decision
p.value <- a_summ[[1]][[1,"Pr(>F)"]]
p.value
ifelse(result$p.value > alpha,"Failed to reject Null Hypothesis","Reject Null Hypothesis")
# 7.e
# Atleast one mean is different from others
```

## Task 8.(Section 12- 3. 10. Increasing Plant Growth)    

A gardening company is testing new ways to improve plant growth. Twelve plants are randomly selected and exposed to a combination of two factors, a “Grow-light” in two different strengths and a plant food supplement with different mineral supplements. After a number of days, the plants are measured for growth, and the results (in inches) are put into the appropriate boxes.      
```{r , echo=FALSE , warning=FALSE, message=FALSE}
df <- data.frame(Plant_Food_Type  = c("Grow-light 1","Grow-light 2"),
                Plant_Food_A  = c(c(9.2,9.4,8.9),c(8.5,9.2,8.9)),
                Plant_Food_B  = c(c(7.1,7.2,8.5),c(5.5,5.8,7.6)))

print(kable(df))
```
Can an interaction between the two factors be concluded? Is there a difference in mean growth 
with respect to light? With respect to plant food? Use alpha = 0.05.

```{r }
# .Method:  Two Way ANOVA Test

# Set significance level
alpha <- 0.05

A_1 <- data.frame('growth' = c(9.2, 9.4, 8.9),
                      'light' = rep(1,3),
                      'plant_food' = rep(1,3),
                      stringsAsFactors = FALSE)

B_1 <- data.frame('growth' = c(7.1, 7.2, 8.5),
                      'light' = rep(1,3),
                      'plant_food' = rep(2,3),
                      stringsAsFactors = FALSE)

A_2 <- data.frame('growth' = c(8.5, 9.2, 8.9),
                      'light' = rep(2,3),
                      'plant_food' = rep(1,3),
                      stringsAsFactors = FALSE)

B_2 <- data.frame('growth' = c(5.5, 5.8, 7.6),
                      'light' = rep(2,3),
                      'plant_food' = rep(2,3),
                      stringsAsFactors = FALSE)

# Joining Data Frames
growth_df <- rbind(A_1,A_2,B_1,B_2)

# (Plot:1) Two-Way Interaction Plot between Density and Fertilizer
interaction.plot(x.factor = round(growth_df$growth,2), trace.factor = growth_df$light,response = growth_df$plant_food, fun = mean, type = "b", legend = TRUE,xlab = "Growth", ylab="Plant Food Type",pch=c(1,19),col = c("#00AFBB", "#E7B800"),trace.label="Light Type")

#  Interaction between Grow Light and Plant Food was noted

# 8.a
# Assuming Expected Frequencies are equal
# Hypotheses
# H1(0): There is no difference in the means of impact of light
# H1(1): Distribution of Frequency is not as stated in null hypothesis (H1(0))
# H2(0): There is no difference in the means of impact of plant food
# H2(1): Distribution of Frequency is not as stated in null hypothesis (H2(0))
# H3(0): There is no interaction between factors impact of light and plant food
# H3(1): Distribution of Frequency is not as stated in null hypothesis (H3(0))

# Factoring Fertilizer and Density
growth_df$light <- as.factor(growth_df$light)
growth_df$plant_food <- as.factor(growth_df$plant_food)

# Compute Two Way ANOVA Test
res.aov2 <- aov(growth ~ light + plant_food, data = growth_df)
a_summ <- summary(res.aov2)

# 8.b 
# (Critical Value) for Light
qf(1-alpha,a_summ[[1]][1,1],a_summ[[1]][2,1])
# (Critical Value) for Plant Food
qf(1-alpha,a_summ[[1]][2,1],a_summ[[1]][2,1])


# 8.c 
# (Test Value) for Light
F.value <- a_summ[[1]][[1,"F value"]]
F.value
# (Test Value) for Plant Food
F.value <- a_summ[[1]][[2,"F value"]]
F.value
# Summary
a_summ

#8.d
# Compare p-value and alpha to make decision

# Hypothesis 1 
p.value <- a_summ[[1]][[2,"Pr(>F)"]]
p.value
ifelse(result$p.value > alpha,"Failed to reject Null Hypothesis","Reject Null Hypothesis")
# Means of Plant Food is Not Equal

# Hypothesis 2 
p.value <- a_summ[[1]][[1,"Pr(>F)"]]
p.value
ifelse(result$p.value > alpha,"Failed to reject Null Hypothesis","Reject Null Hypothesis")
# Mean of Grow Light Not Equal

# Hypothesis 3
a_summ
# 8.e
# It can be seen that Plant food had more significant on growth, very little significance
# of grow light was observed
# Also, interaction between Plant Food Type A and Grow Lights where observed with highest
# growth for Plant Food A and Light 1
```


## Task 9 (baseball.csv and crop_data.csv)

```{r }
# 9.1 Importing baseball.csv
baseball_data <- read.csv(
  "/Users/davidjjohn97/Northeastern/Quarter_2/Term_3/Intermediate_Analytics/baseball.csv",
  header = TRUE)
df1<-baseball_data

# 9.2 Exploratory Data Analysis
## Head,Tail and Structure of Dataset
head(df1)
tail(df1)
str(df1)

## Summary of Dataset
summary(df1)

## Describing the Dataset
describe(df1)

## Visualizations

### Filtering Top 10 Winners
df_sorted <- df1[order(-df1$W),]
top_10_win <- head(df_sorted,10)

last_10_win <- df1[order(df1$W),]
last_10_win <-head(last_10_win,10)
### Histogram of Number of Games
hist(df1$Year, main="(Plot:2) Yearly Number of Games", xlab = "Quinquennial Groups")

### Scatterplot of Runs Scored Vs Runs Allowed of top 10 Teams
plot(top_10_win$RS ~ top_10_win$RA,data=top_10_win,pch=20,xlab="Runs Allowed",ylab = "Runs Scored",xlim = c(0,1000),ylim = c(0,1000),col="DarkBlue",main="(Plot:3) Scatterplot of Runs Scored Vs Runs Allowed of top 10 Teams")
abline(lm(top_10_win$RS~top_10_win$RA, data = top_10_win),col="green")

### Scatterplot of Runs Scored Vs Runs Allowed of top 10 Teams
plot(last_10_win$RS ~ last_10_win$RA,data=last_10_win,pch=20,xlab="Runs Allowed",ylab = "Runs Scored",col="DarkBlue",main="(Plot:4) Scatterplot of Runs Scored Vs Runs Allowed of bottom 10 Teams",xlim = c(0,1000),ylim = c(0,1000))
abline(lm(last_10_win$RS~last_10_win$RA, data = last_10_win),col="red")

### Imputing numeric columns using package imputeTS
df3 <- na_mean(df1)
df3<-select_if(df1, is.numeric)
df3 = subset(df3, select = -c(RankSeason,RankPlayoffs,G,OOBP,OSLG) )
matrix_format <- cor(df3)
matrix_format
### (Plot:5) Correlation Plot
corrplot(matrix_format,tl.cex=0.65,diag=TRUE)
```
Insights Drawn:

* Higher Number of Matches where played between 2000 -2010

* Scoring around 800 runs while allowing only 600 runs shows more odds of winning 

* Scoring only 600 runs while allowing around 800 runs shows more odds of losing

* On-base Percentage, Slugging Percentage and Batting Average had high correlation with the Run Scored.Batting Average had high correlation with the On-base Percentage and Slugging Percentage, hence statistically validating the higher score for points scored in this manner.

* The decade 2000 had highest number of wins

```{r }
# 9.3 Chi-Square Goodness of Fit to test difference in number of wins over the decades

# Extract decade from year 
df1$Decade <- df1$Year - (df1$Year %% 10) 
 

# Create a wins table by summing the wins by decade 
wins <- df1 %>% group_by(Decade) %>%summarize(wins = sum(W)) %>% tibble::as_tibble()
wins


# Plotting Result Table
par(mar=c(10, 5, 4, 1))
bp <-barplot(wins$wins,main="(Plot:6) Wins over the Decade",xlab="Decade",ylab="Wins",
            col = "lightblue",names.arg=c("1960", "1970s", "1980","1990","2000","2010"))
text(bp, 0, round(wins$wins, 1),cex=1,pos=3)

# Method:  Chi-Square Goodness of Fit Test
# Assuming Expected Frequencies are equal
# 9.3.a (Hypotheses)
# H0: Frequency of wins are Equal
# H1: Distribution of Frequency is not as stated in null hypothesis

# Set significance level
alpha <- 0.05
# Assuming frequency of win is equal in each decade,
# probability of win for each decade will be  probabilities of general population
prob <- c(1/6,1/6,1/6,1/6,1/6,1/6)
# Vector of observed values of sample
observed <- wins$wins

result <- chisq.test(x= observed, p= prob)
# Test Statistic
result
# 9.3..b (Critical Value)
critical_value = qchisq(p=alpha, df=result$parameter, lower.tail=FALSE)
critical_value
# 9.3..c (Test Value)
result$statistic 
# 9.3..d
# Compare p-value and alpha to make decision
ifelse(result$p.value > alpha,"Failed to reject Null Hypothesis","Reject Null Hypothesis")
# 9.3.e
# Compare critical value and test value to make decision
ifelse(result$statistic < critical_value | (var(critical_value,result$statistic) > 100),
       "Failed to reject Null Hypothesis","Reject Null Hypothesis")

# Frequency of wins are NOT Equal, Null Hypothesis is rejected comparing p-value
# and alpha as well as critical value and test value

# 9.4 Importing crop_data.csv
crop_data <- read.csv(
  "/Users/davidjjohn97/Northeastern/Quarter_2/Term_3/Intermediate_Analytics/crop_data.csv",
  header = TRUE)

# 9.5 Two-Way Anova Test to check impact of fertilizer and density on yield
df2<-crop_data

# (Plot:7) Two-Way Interaction Plot between Density and Fertilizer
interaction.plot(x.factor = round(df2$yield,2), trace.factor = df2$density,
                 response = df2$fertilizer, fun = mean, type = "b", legend = TRUE,
                 xlab = "Yield", ylab="Fertilizer",pch=c(1,19),
                 col = c("#00AFBB", "#E7B800"),
                 trace.label="Density")

#  Interaction between fertilizer and density was noted with
#  highest yield for Fertilizer 3 with Density 2



# Method: Two Way ANOVA Test
# Assuming Expected Frequencies are equal
# Hypotheses
# H1(0): There is no difference in the means of impact of fertilizer
# H1(1): Distribution of Frequency is not as stated in null hypothesis (H1(0))
# H2(0): There is no difference in the means of impact of density
# H2(1): Distribution of Frequency is not as stated in null hypothesis (H2(0))
# H3(0): There is no interaction between factors impact of fertilizer and density
# H3(1): Distribution of Frequency is not as stated in null hypothesis (H3(0))

# Factoring Fertilizer and Density
df2$density <- as.factor(df2$density)
df2$fertilizer <- as.factor(df2$fertilizer)

# Compute Two Way ANOVA Test
res.aov2 <- aov(yield ~ density + fertilizer, data = df2)
a_summ <- summary(res.aov2)


# Compare p-value and alpha to make decision

# Hypothesis 1 
p.value <- a_summ[[1]][[2,"Pr(>F)"]]
p.value
ifelse(result$p.value > alpha,"Failed to reject Null Hypothesis","Reject Null Hypothesis")
# Means of Impact of Fertilizer is Not Equal

# Hypothesis 2 
p.value <- a_summ[[1]][[1,"Pr(>F)"]]
p.value
ifelse(result$p.value > alpha,"Failed to reject Null Hypothesis","Reject Null Hypothesis")
# Mean of Density Not Equal

# Hypothesis 3
a_summ
# It can be seen that the two main effects (fertilizer and density) are statistically significant,
# as well as their interaction.
```

\newpage

# Conclusion

* Task 1 : Distribution of Blood Groups in the sample was same as that of the population

* Task 2 : Distribution of observed results differ from the government’s statistics

* Task 3 : Movie Admissions where found to be dependent on the Ethinicity

* Task 4 : Ranks of women in military are dependent of branch

* Task 5 : It was concluded that a difference in mean sodium amounts exists among condiments, cereals, and desserts

* Task 6 : It was concluded that a difference in mean sales amounts exists among chocolates, cereals, and candy

* Task 7 : It was concluded that a difference in mean expenditures (in dollars) per pupil for states in three sections of the country

* Task 8 : It can be seen that Plant food had more significant on growth, very little significance of grow light was observed.Also, interaction between Plant Food Type A and Grow Lights where observed with highest growth for Plant Food A and Light 1.

* Task 9 : Baseball Dataset : Frequency of wins are NOT Equal, Null Hypothesis is rejected
            comparing p-value and alpha as well as critical value and test value;
           Crop Dataset : Neither Mean Impact of Fertilizer neither Mean Density are Equal,
            however effects of (fertilizer and density) are statistically significant on yield;


\newpage

# References

* How Hypothesis Testing Works. (2022). Retrieved 20 January 2022, from https://www.investopedia.com/terms/h/

* The Chi-Square Test. (2022). Retrieved 20 January 2022, from https://www.jmp.com/en_us/statistics-knowledge-portal/chi-square-test.html

* ANOVA Test: Definition, Types, Examples, SPSS. (2022). Retrieved 20 January 2022, from https://www.statisticshowto.com/probability-and-statistics/hypothesis-testing/anova/#:~:text=An%20ANOVA%

* Zach, V. (2020). How to Find the Chi-Square Critical Value in R - Statology. Retrieved 23 January 2022, from https://www.statology.org/chi-square-critical-value-r/

* S.3.1 Hypothesis Testing (Critical Value Approach). (2022). Retrieved 23 January 2022, from https://online.stat.psu.edu/statprogram/b

* How to find the critical value of F for one-way ANOVA in R?. (2022). Retrieved 23 January 2022, from https://www.tutorialspoint.com/how-to-find-the-critical-value-of-f-for-one-way-anova-in-r


